---
title: "Skitouring"
author: "Kaspar Rufibach"
date: '`r Sys.setlocale("LC_TIME", "C"); format(Sys.Date(), "%d %b %Y")`'
output:
  html_document:
    keep_md: true
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document: 
    toc: true
    toc_depth: 2
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
## load packages
packs.html <- c("knitr", "pander", "reporttools", "dplyr", "lubridate", "readxl")
for (i in 1:length(packs.html)){library(packs.html[i], character.only = TRUE)}

knitr::opts_chunk$set(echo = TRUE)

path <- "C:/rufibach/01_personal/05_Sport/40_reporting/"
path <- paste(getwd(), "/", sep = "")

## =================================================================
## input data
## =================================================================
touren <- read_excel(paste(path, "data/touren.xls", sep = ""))

# Hoehenmeter
hm <- read_excel(paste(path, "data/hm.xlsx", sep = ""))
hm$mum <- with(hm, pmax(mum_prov, mum_def, na.rm = TRUE))

# merge the two and add time in hours
touren <- (inner_join(touren, hm[, c("ort", "mum")], by = c("start" = "ort")) %>% rename(start_hm = mum) %>% inner_join(hm[, c("ort", "mum")], by = c("end" = "ort")) %>% rename(end_hm = mum) %>% mutate(time_h = as.numeric(hms(time, quiet = TRUE)) / 3600))

# add Hm / h
touren <- (touren %>% mutate(hm_diff = end_hm - start_hm, hm_h = hm_diff / time_h))

# add season
cuts <- parse_date_time(x = "2000-06-30", orders = "ymd") + years(x = seq.int(from = 0, to = 50, by = 1))
touren$season <- cut.POSIXt(x = parse_date_time(x = touren$date, orders = "ymd"), breaks = cuts, 
                         labels = c(paste("Winter 0", 0:8, "/0", 1:9, sep = ""), 
                                    "Winter 09/10", paste("Winter ", 10:49, "/", 11:50, sep = "")))

# skitouren
all <- (touren %>% mutate(date = as.Date(date, origin = "1899-12-30"), skihalbtag = NULL, skihalbtag_fam = NULL, begleitung1 = NULL, begleitung2 = NULL, begleitung3 = NULL, region = NULL, kommentar = NULL) %>% filter(sport == "Skitour") %>% arrange(end, time) )
all <- as.data.frame(all)
all[is.na(all)] <- ""

# skitage
skitage1 <- touren %>% mutate(date = as.Date(date, origin = "1899-12-30"), skihalbtag = as.numeric(skihalbtag), skihalbtag_fam = as.numeric(skihalbtag_fam)) %>% filter(sport == "Skifahren") %>% arrange(end, time) 

```

# Summary over years

List by far not complete. Multiple tours within same month are pooled and frequency is given.

```{r, echo = FALSE, results = 'asis', message = FALSE, eval = FALSE}
with(subset(touren, sport == "Skitour"), pander(addmargins(table(as.character(end), year(date))), digits = 1))
```

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
par(mfrow = c(1, 1), las = 2, mar = c(4, 8, 1, 1))

ski <- (all %>% select(start, end, date, time))
tour <- unique(ski$end)

# x-axis
xs <- ymd_hms(paste(min(year(ski$date)):max(year(ski$date) + 1), "0101 00:00:00", sep = ""))
plot(xs, rep(0, length(xs)), type = "n", ylim = c(0, length(tour)), xaxt = "n", yaxt = "n", xlab = "", ylab = "")
axis(1, at = xs, labels = year(xs))
axis(2, at = length(tour):1, labels = tour)
abline(h = 1:length(tour), lty = 2, col = grey(0.5))
offset <- 0.2

for (i in 1:length(tour)){
  dat.i <- (ski %>% filter(end == tour[length(tour) - i + 1]) %>% arrange(date) %>% group_by("w" = paste(year(date), month(date), sep = "")) %>% add_tally() %>% slice(1))
  xs.i <- ymd_hms(paste(as.character(dat.i$date), " 00:00:00", sep = ""))
  text(xs.i, rep(i, nrow(dat.i)), dat.i$n)
}
```

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
par(las = 2, mar = c(8, 4.5, 1, 1))

# now pick first season with entry and last season
dat <- table(all$season)
i1 <- which(names(dat) == unique(sort(droplevels(all$season)))[1])
i2 <- which(names(dat) == tail(unique(sort(droplevels(all$season))), 1))
dat <- dat[i1:i2]

plot(dat, type = "b", col = 2, xlab = "", ylab = "Anzahl Skitouren / Skihalbtage")
abline(h = 1:100, lty = 2, col = grey(0.5))

# Skihalbtage 
ski1 <- (skitage1 %>% group_by(season) %>% summarize(sum = sum(skihalbtag, na.rm = TRUE), .groups = 'drop'))
dat1 <- dat
dat1[1:length(dat1)] <- NA
dat1[names(dat1) %in% ski1$season] <- ski1$sum
lines(1:length(dat1), dat1, col = 3, type = "b", pch = 19)

# Familien-Skihalbtage 
ski2 <- (skitage1 %>% group_by(season) %>% summarize(sum = sum(skihalbtag_fam, na.rm = TRUE), .groups = 'drop'))
dat2 <- dat
dat2[1:length(dat2)] <- NA
dat2[names(dat2) %in% ski1$season] <- ski2$sum
lines(1:length(dat2), dat2, col = 4, type = "b", pch = 19)
```

# Days on ski in total

```{r echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
s <- 21
f1 <- lapply(with(touren, split(skihalbtag, season)), sum, na.rm = TRUE)[[s]]
f2 <- lapply(with(touren, split(skihalbtag_fam, season)), sum, na.rm = TRUE)[[s]]
f3 <- lapply(with(touren, split(sport == "Skitour", season)), sum, na.rm = TRUE)[[s]]

f_tab <- data.frame(cbind(c("Ski half-day", "Ski half-day family", "Skitour"), c(f1, f2, f3)))
colnames(f_tab) <- c("Type", "frequency")
pander(f_tab)

# days on ski (double counting not possible)
f4 <- with(touren, data.frame(date, skihalbtag, skihalbtag_fam, tour = as.numeric(sport == "Skitour"), season))
f5 <- f4[f4$season == levels(with(touren, season))[s], ]
f5$any <- with(f5, pmax(skihalbtag, skihalbtag_fam, tour, na.rm = TRUE))
f5_dat <- f5[f5$any == 1, "date"]
skidays <- length(unique(f5_dat))
```

Total days on skis: `r skidays`.

# Listings

This section provides listings, chronologically and by time. For listings by time, tours with missing time are omitted.

```{r echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}

# how to do this with child documents: https://gist.github.com/rmoff/a043676a2f084b81a434

out <- NULL
for (i in 1:length(tour)){
  out <- c(out, knit_expand('tour_summary.Rmd'))
}

cat(knit(text = unlist(paste(out, collapse = '\n')), quiet = TRUE))
```

# Elevation gain per hour

This section provides a list of tours be elevation gain per hour, sorted in decreasing order.

```{r, echo = FALSE, results = 'asis', message = FALSE, eval = TRUE}
t <- (touren %>% filter(is.na(hm_h) == FALSE & sport == "Skitour") %>% select(start, end, date, hm_diff, time, hm_h) %>% arrange(desc(hm_h)) %>% mutate(start = substr(start, 1, 23)))
pander(t)
```


