---
title: "Hikes, runs, and rides"
author: "Kaspar Rufibach"
date: '`r Sys.setlocale("LC_TIME", "C"); format(Sys.Date(), "%d %b %Y")`'
output:
  html_document:
    keep_md: true
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document: 
    toc: true
    toc_depth: 2
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
## load packages
packs.html <- c("knitr", "pander", "reporttools", "dplyr", "readxl", "lubridate")
for (i in 1:length(packs.html)){library(packs.html[i], character.only = TRUE)}

knitr::opts_chunk$set(echo = TRUE)

path <- paste(getwd(), "/", sep = "")

source(paste(path, "/functions/input_touren.r", sep = ""))
source(paste(path, "/functions/hm_print.r", sep = ""))

## =================================================================
## input data
## =================================================================
touren <- input_touren(path)

all0 <- (touren %>% mutate(begleitung1 = NULL, begleitung2 = NULL, region = NULL, kommentar = NULL) %>% arrange(sport, end, time))
```

# Bike

## Listings

For listings by time, tours with missing time are omitted.

```{r bike, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}

# how to do this with child documents: https://gist.github.com/rmoff/a043676a2f084b81a434

all <- (all0 %>% filter(sport == "Bike"))
tour <- unique(all$end)
out1 <- NULL
for (i in 1:length(tour)){
  out1 <- c(out1, knit_expand('tour_summary.Rmd'))
}

cat(knit(text = unlist(paste(out1, collapse = '\n')), quiet = TRUE))
```

## Elevation gain per hour

This section provides a list of tours be elevation gain per hour, sorted in decreasing order.

```{r bike2, echo = FALSE, results = 'asis', message = FALSE, eval = TRUE}
hm_print(touren %>% filter(is.na(hm_h) == FALSE & sport == "Bike"))
```

# Run

## Listings

```{r run, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}

# how to do this with child documents: https://gist.github.com/rmoff/a043676a2f084b81a434

all <- (all0 %>% filter(sport == "Run"))
tour <- unique(all$end)
out2 <- NULL
for (i in 1:length(tour)){
  out2 <- c(out2, knit_expand('tour_summary.Rmd'))
}

cat(knit(text = unlist(paste(out2, collapse = '\n')), quiet = TRUE))
```

## Elevation gain per hour

This section provides a list of tours be elevation gain per hour, sorted in decreasing order.

```{r run2, echo = FALSE, results = 'asis', message = FALSE, eval = TRUE}
hm_print(touren %>% filter(is.na(hm_h) == FALSE & sport == "Run"))
```

# Bike and run

## Listings

```{r hike, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}

# how to do this with child documents: https://gist.github.com/rmoff/a043676a2f084b81a434

all <- (all0 %>% filter(sport == "Bike and run"))
tour <- unique(all$end)
out3 <- NULL
for (i in 1:length(tour)){
  out3 <- c(out3, knit_expand('tour_summary.Rmd'))
}

cat(knit(text = unlist(paste(out3, collapse = '\n')), quiet = TRUE))
```

## Elevation gain per hour

This section provides a list of tours be elevation gain per hour, sorted in decreasing order.

```{r hike2, echo = FALSE, results = 'asis', message = FALSE, eval = TRUE}
hm_print(touren %>% filter(is.na(hm_h) == FALSE & sport == "Bike and run"))
```

# Hike

## Listings

```{r bikeandrun, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}

# how to do this with child documents: https://gist.github.com/rmoff/a043676a2f084b81a434

all <- (all0 %>% filter(sport == "Hike"))
tour <- unique(all$end)
out4 <- NULL
for (i in 1:length(tour)){
  out4 <- c(out4, knit_expand('tour_summary.Rmd'))
}

cat(knit(text = unlist(paste(out4, collapse = '\n')), quiet = TRUE))
```

## Elevation gain per hour

This section provides a list of tours be elevation gain per hour, sorted in decreasing order.

```{r bikeandrun2, echo = FALSE, results = 'asis', message = FALSE, eval = TRUE}
hm_print(touren %>% filter(is.na(hm_h) == FALSE & sport == "Hike"))
```