---
title: "Yearly summaries"
author: "Kaspar Rufibach"
date: '`r Sys.setlocale("LC_TIME", "C"); format(Sys.Date(), "%d %b %Y")`, as per 23rd May 2020'
output:
  html_document:
    keep_md: true
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  github_document: 
    toc: true
    toc_depth: 2
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
## load packages
packs.html <- c("knitr", "pander", "reporttools", "dplyr", "readxl", "ggplot2")
for (i in 1:length(packs.html)){library(packs.html[i], character.only = TRUE)}

knitr::opts_chunk$set(echo = TRUE)

path <- paste(getwd(), "/", sep = "")

source(paste(path, "functions/plotSummary.r", sep = ""))

## =================================================================
## input 
## =================================================================
sportKR <- as.data.frame(read_excel(paste(path, "data/zsf.xlsx", sep = ""), col_types = NULL))
sportKR <- (sportKR %>% mutate("Sport" = as.factor(Sport), "Einheit" = as.factor(Einheit)))

lastcol <- ncol(sportKR)

# define which years to include
include <- sportKR[sportKR[, "Sport"] == "Include", 3:ncol(sportKR)]
include <- as.numeric(include) * (3:ncol(sportKR))
include <- include[include > 0] 

# collapse certain disciplines
sportKR[sportKR[, "Sport"] == "Bike", 3:ncol(sportKR)] <- sportKR[sportKR[, "Sport"] == "Bike", 3:ncol(sportKR)] + sportKR[sportKR[, "Sport"] == "Velo", 3:ncol(sportKR)] 

sportKR[sportKR[, "Sport"] == "Running", 3:ncol(sportKR)] <- sportKR[sportKR[, "Sport"] == "Running", 3:ncol(sportKR)] + sportKR[sportKR[, "Sport"] == "Trail Running", 3:ncol(sportKR)] 

# remove unused rows
sportKR <- (sportKR %>% filter(!(Sport %in% c("Include", "Velo", "Trail Running"))))

# linearly extrapolate last column to full year
toda <- Sys.Date()
first <- as.Date(paste(substr(toda, 1, 4), "-01-01", sep = ""))
frac <- as.numeric((toda - first) / 365.25)
#sportKR[, lastcol] <- sportKR[, lastcol] / frac
#include <- c(include, tail(include, 1) + 1)

# add extrapolation, for simplicity give it next year as label
sportKR[, as.character(as.numeric(tail(colnames(sportKR), 1)) + 1)] <- sportKR[, lastcol] / frac
dat <- sportKR[, 3:ncol(sportKR)]
include <- c(include, tail(include, 1) + 1)

# only use years defined by "include"
sportKR <- sportKR[, c(1, 2, include)]
dat <- dat[, include - 2]

# choose scaling
skal <- c(1, 52)
skal.nam <- c(" per year", " per week")

# monthly ascent
ascent <- as.data.frame(read_excel(paste(path, "data/ascentmonthly.xlsx", sep = ""), col_types = NULL))
```

<!-- Numbers labelled `r tail(colnames(dat), 1)` are linearly extrapolated from `r toda` to the full year of `r tail(colnames(dat), 2)[1]`. -->

# Yearly summaries {.tabset .tabset-fade .tabset-pills}

The year 2021 always refers to the current status of 2020 linearly extrapolated to a full year.

## Kilometer

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
par(mfrow = c(1, 1), mar = c(3.5, 4.5, 3, 1), las = 2)
km <- subset(sportKR, Einheit == "Km")
xs <- as.numeric(colnames(km)[3:ncol(km)])
ys <- km[, 3:ncol(km)]
plot(0, 0, type = "n", xlim = range(xs), ylim = c(0, max(ys)), xlab = "", ylab = "Kilometer", 
     main = "Kilometer")
for (i in 1:nrow(km)){lines(xs, ys[i, ], type = "b", col = i + 1, lwd = 2)}
legend("bottomleft", legend = km$Sport, lty = 1, col = 2:(nrow(km) + 1), bty = "n", lwd = 2)
```

## Ascent

```{r, echo = FALSE, results = 'asis', message = FALSE}
km <- subset(sportKR, Einheit == "Hm")
xs <- as.numeric(colnames(km)[3:ncol(km)])
ys <- km[, 3:ncol(km)]
ystot <- apply(ys, 2, sum)
ystot <- rbind(ystot, ystot / 52, ystot / 365)
rownames(ystot) <- c("Total", "per week", "per day")
pander(ystot) 
```

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
par(mfrow = c(1, 1), mar = c(3.5, 5, 3, 1), las = 2)
plot(0, 0, type = "n", xlim = range(xs), ylim = c(0, max(ys)), xlab = "", ylab = "", 
     main = "Ascent")
for (i in 1:nrow(km)){lines(xs, ys[i, ], type = "b", col = i + 1, lwd = 2)}
legend("topleft", legend = km$Sport, lty = 1, col = 2:(nrow(km) + 1), bty = "n", lwd = 2)

par(las = 0)
mtext(text = "Ascent", side = 2, line = 4, adj = 0.5)
```

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
df <- rbind(cbind(xs, t(ys)[, 1]), cbind(xs, t(ys)[, 2]), cbind(xs, t(ys)[, 3]), cbind(xs, t(ys)[, 4]))
df <- data.frame(df, rep(as.character(km[, "Sport"]), each = length(xs)))
colnames(df) <- c("Jahr", "Hm", "Disziplin")
ggplot(df, aes(x = Jahr, y = Hm, fill = Disziplin)) + geom_area(position = 'stack')
```

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
par(mfrow = c(1, 1), mar = c(6, 5, 3, 1), las = 2)
plot(0, 0, type = "n", xlim = c(1, 12), ylim = c(0, 20000), xlab = "", ylab = "", main = "Ascent", xaxt = "n")
for (i in 2:ncol(ascent)){lines(1:12, ascent[, i], type = "b", col = i, lwd = 2)}
axis(1, at = 1:12, labels = ascent[, "month"])
par(las = 0)
mtext(text = "Ascent (Hm)", side = 2, line = 4, adj = 0.5)
legend("bottomleft", legend = paste(colnames(ascent)[-1], ": ", apply(ascent[, -1], 2, sum, na.rm = TRUE), sep = ""), lty = 1, col = 2:ncol(ascent), bty = "n", lwd = 2)
```

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
par(mfrow = c(1, 1), mar = c(6, 5, 5, 1), las = 2)
plot(0, 0, type = "n", xlim = c(1, 12), ylim = c(0, 100000), xlab = "", ylab = "", main = "Cumulative ascent", xaxt = "n", yaxt = "n")
for (i in 2:ncol(ascent)){
  tmp1 <- ascent[, i]
  tmp2 <- cumsum(na.omit(tmp1))
  if (length(tmp2) < 12){tmp2 <- c(tmp2, rep(NA, sum(is.na(tmp1))))}
  
  lines(1:12, tmp2, type = "b", col = i, lwd = 2)
  }
axis(1, at = 1:12, labels = ascent[, "month"])
axis(2, at = seq(0, 10 ^ 5, by = 10 ^ 4), labels = format(seq(0, 10 ^ 5, by = 10 ^ 4), scientific = FALSE))
par(las = 0)
mtext(text = "Cumulative ascent (Hm)", side = 2, line = 4, adj = 0.5)
legend("topleft", legend = paste(colnames(ascent)[-1], ": ", apply(ascent[, -1], 2, sum, na.rm = TRUE), sep = ""), lty = 1, col = 2:ncol(ascent), bty = "n", lwd = 2)
gap <- (cumsum(na.omit(ascent[, ncol(ascent)])) - cumsum(na.omit(ascent[, ncol(ascent) - 1])))[tail(which(is.na(ascent[, ncol(ascent)]) == FALSE), 1)]
mtext(paste("Delta to last year, last month = ", gap, " Hm", sep = ""), side = 3, line = 0.5)
```


## Time

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
par(mfrow = c(1, 1), mar = c(3.5, 4.5, 3, 1), las = 2)
km <- subset(sportKR, Einheit == "Time")
xs <- as.numeric(colnames(km)[3:ncol(km)])
ys <- km[, 3:ncol(km)]
plot(0, 0, type = "n", xlim = range(xs), ylim = c(0, max(ys)), xlab = "", ylab = "Stunden", 
     main = "Stunden")
for (i in 1:nrow(km)){lines(xs, ys[i, ], type = "b", col = i + 1, lwd = 2)}
legend("bottomleft", legend = km$Sport, lty = 1, col = 2:(nrow(km) + 1), bty = "n", lwd = 2)
```

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
df <- rbind(cbind(xs, t(ys)[, 1]), cbind(xs, t(ys)[, 2]), cbind(xs, t(ys)[, 3]))
df <- data.frame(df, rep(as.character(km[, "Sport"]), each = length(xs)))
colnames(df) <- c("Jahr", "Time", "Disziplin")
ggplot(df, aes(x = Jahr, y = Time, fill = Disziplin)) + geom_area(position = 'stack')
```

## Places

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 7}
par(mfrow = c(1, 1), mar = c(3.5, 4.5, 3, 1), las = 2)
km <- subset(sportKR, Einheit == "Number" & (Sport %in% c("Exersuisse")) == FALSE)
xs <- as.numeric(colnames(km)[3:ncol(km)])
ys <- km[, 3:ncol(km)]
plot(0, 0, type = "n", xlim = range(xs), ylim = c(0, max(ys)), xlab = "", ylab = "number", 
     main = "Number")
for (i in 1:nrow(km)){lines(xs, ys[i, ], type = "b", col = i + 1, lwd = 2)}
legend("topleft", legend = km$Sport, lty = 1, col = 2:(nrow(km) + 1), bty = "n", lwd = 2)
```

# Summary over winter seasons (not years)

## Skiing

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 7}
par(mfrow = c(1, 1), mar = c(6, 4.5, 3, 1), las = 2)
km <- subset(sportKR, Einheit == "Number" & (Sport %in% c("Skiing", "Skitouring")) == TRUE)
xs <- as.numeric(colnames(km)[3:ncol(km)])
ys <- km[, 3:ncol(km)]
plot(0, 0, type = "n", xlim = range(xs), ylim = c(0, max(ys)), xlab = "", ylab = "number", 
     main = "Number", xaxt = "n")
axis(1, at = xs, labels = paste(xs - 1, " / ", xs, sep = ""))
for (i in 1:nrow(km)){lines(xs, ys[i, ], type = "b", col = i + 1, lwd = 2)}
legend(2013.5, 14, legend = km$Sport, lty = 1, col = 2:(nrow(km) + 1), bty = "n", lwd = 2)
```

# Activity per year {.tabset .tabset-fade .tabset-pills}

## Bike

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
s <- 1
for (i in 1:3){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

## Run

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
for (i in 4:6){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

## Hiking

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
for (i in 7:9){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

## Exersuisse

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
for (i in 10){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```


## Orte

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
for (i in 13:nrow(dat)){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

# Activity per week {.tabset .tabset-fade .tabset-pills}

## Bike

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
s <- 2
par(mar = c(4, 4, 4, 1), las = 1)
for (i in 1:3){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

## Run

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
for (i in 4:6){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

## Hiking

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
for (i in 7:9){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

## Exersuisse

```{r, echo = FALSE, fig.cap = "", fig.align = "center", fig.width = 7, fig.height = 5}
for (i in 10){plotSummary(dat = dat, sportKR = sportKR, i = i, s = s)}
```

